First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2017-04-20 eg6-letters/coherent/dvcs_he4_prl.tex
eg6-letters/coherent:  implementing adhoc comments, Part C.  Remainder is restructure of experimental setup (paragraphs starting at lines 82 and 112) and multi-gamma explanation (line 137)

One must overcome two major experimental challenges in measuring coherent 
nuclear DVCS, $eA\rightarrow e^\prime A^\prime\gamma$. First, the cross section 
of coherent scattering is suppressed due to the nuclear form factor, and 
second, the recoil nucleus ($A^\prime$) must be detected to ensure coherence.  
Figure \ref{fig:diags} illustrates the handbag diagram for coherent DVCS on 
$^4$He, where four-vectors of the electrons, photons, and $^4$He are denoted by
$k/k^\prime$, $q/q^\prime$, and $p/p^\prime$ respectively. 
For large virtual photon's 4-momentum squared, $Q^2=-(k-k')^{2}$, and small squared momentum transfer, $t=(p-p')^{2}$, the DVCS handbag 
diagram can be factorized into two parts \cite{Freund_Collins,Ji_Osborne}. The 
hard part includes the photon-quark interaction and is calculable in 
perturbative QCD. The non-perturbative part is parametrized in terms of GPDs, 
which embed the partonic structure of the hadron. The GPDs depend on the three 
variables $x$, $\xi$, and $t$ as introduced in Fig.~\ref{fig:diags}. The 
parameter $\xi$ relates to the Bjorken variable $x_{B}$: $\xi\approx 
{{x_B}\over{2-x_B}}$, where $x_B=\frac{Q^2}{2M\nu}$, $\nu=k^0-k^{0\prime}$, and 
$M$ is the proton mass. The parameter $x$ is the quark's internal loop momentum 
fraction and cannot be accessed experimentally. We in fact measure 
Compton form factors (CFF) \cite{Guidal:2013rya}, which are
complex amplitudes defined as:

@@

The $^4$He nucleus is an 
ideal experimental target in this regard, as it is light enough to be detected 
in such a setting, while it is subject to significant nuclear effects 
\cite{JSeely} and has rather high density. A helium target leads to another 
important advantage, as the number of GPDs defined for a hadron depends on its 
spin. The structure of a spin-zero nucleus, such as $^4$He, is parametrized by 
only one chiral even GPD ($H_{A}(x,\xi,t)$) at leading twist, while 4 GPDs 
arise in the nucleon case. This significantly simplifies the interpretation of 
the data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

The CEBAF large acceptance spectrometer in Hall-B at 
Jefferson Lab has been previously used successfully for DVCS measurements on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}.

In order to measure DVCS on $^4$He with a fully exclusive final state, we built 
a specialized radial time projection chamber (RTPC) for the detection and 
identification of low-energy recoiling light nuclei.

%experimental setup

The experiment E08-024 took place in Hall-B at Jefferson Lab in 2009 using the 
nearly 100\% duty factor, longitudinally polarized electron beam (83$\%$ 
polarization) at an energy of 6.064 GeV. The data were collected over 40 days 
using a 6 atm gaseous $^4$He target placed 64 cm upstream of the nominal center 
of CLAS. For DVCS experiments, the CLAS baseline design \cite{Mecking:2003zu} 
is supplemented with an inner calorimeter (IC) and a solenoid magnet. The IC 
extends the photon detection acceptance of CLAS to polar angles as low as 
4$^{\circ}$. The low-energy M\o{}ller electrons produced in the target form a 
very high rate background that is suppressed by a 5-Tesla solenoid magnet 
placed around the target.

@@

The DVCS and well-known Bethe-Heitler (BH) processes, in which the real 
photon is emitted by the incoming or the outgoing lepton, have the same final 
state and are indistinguishable. The amplitude of electroproduction of real
photons includes a sum of the amplitudes of these two processes. The BH 
amplitude is defined by the target form factors, while the DVCS amplitude is 
a combination of the form factors and GPDs. For our kinematics, the cross 
section of real photon electroproduction is dominated by the BH 
contribution, while the DVCS contribution is very small. Its effect is however
enhanced in the observables sensitive to the interference term, {\it e.g.} 
BSA. The three terms entering the cross section calculation,
the square of the BH and DVCS amplitudes and their interference term, depend on the
azimuthal angle
$\phi$ between the $(e,e^\prime)$ and $(\gamma^*,^4$He$^\prime)$ planes, as
shown for the nucleon in Ref.~\cite{Belitsky:2001ns} and for the spin-zero targets
in Refs.~\cite{Kirchner:2003wt,Belitsky:2008bz}. Based on this work, the BSA 
of a spin-zero hadron can be expressed as
\begin{equation}
\begin{split}
&A_{LU}(\phi) = \\
&\frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( \Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)},
\end{split}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the $^4$He CFF $\mathcal{H}_{A}$ and depend on the GPD $H_A$.  
Explicit expressions of the purely kinematic factors $\alpha_i$ are written in
Ref.~\cite{Belitsky:2008bz} and are functions of Fourier harmonics in the 
azimuthal angle $\phi$, the nuclear form factor $F_A(t)$, $Q^2$, $x_B$,
$x_{A} = \frac{x_{B}M_{N}}{M_{A}}$, $y=\frac{Q^{2}}{2x_{A}M_{A}E_{beam}}$, and 
$\epsilon = \frac{2x_{A}M_{A}}{\sqrt{Q^{2}}}$, where $M_{A}$($M_{N}$) is the $^4$He 
(nucleon) mass. Using the different $\sin(\phi)$ and $\cos(\phi)$ contributions 
in these factors, it is possible to unequivocally extract both imaginary and real 
parts of $\mathcal{H}_{A}$ from the BSA.

In this work, the azimuthal dependence of the beam-spin asymmetry has been studied 
for a wide range of kinematics ($Q^2$ between 1 and 2.2 GeV$^2$/c$^2$, $x_B$ between 0.1 and 0.3, and $-t$ between 0.07 and 0.16 GeV$^2$/c$^2$). We identified two main backgrounds to our 
measurement, accidental coincidences and exclusive coherent $\pi^0$ production.  
The accidentals have particles originating from different events, and we 
estimated their contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all our cuts but with an electron and 
helium originating from different vertices. The $\pi^0$ production can easily
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low energy in the laboratory frame and remains undetected.  
To estimate the effect of this background, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ measured by our 
experiment. We used this generator together with a GEANT3 simulation of our 
detector to estimate the ratio between the number of $\pi^0$ events where the 
two photons are detected and those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data.  Depending on the kinematics, we found contaminations of 
2 to 4\%. The study of systematic uncertainties showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%), where the values are relative and for $A_{LU}$ at $\phi=90^\circ$.
However, added quadratically, the total systematic uncertainty is about 10\%,
which is significantly smaller than statistical  uncertainties in all kinematical bins.

>>>>

One must overcome two major experimental challenges in measuring coherent 
nuclear DVCS, $eA\rightarrow e^\prime A^\prime\gamma$. First, the cross section 
of coherent scattering is suppressed due to the nuclear form factor, and 
second, the intact recoil nucleus ($A^\prime$) must be detected to ensure coherence.  
Figure \ref{fig:diags} illustrates the handbag diagram for coherent DVCS on 
$^4$He, where four-vectors of the electrons, photons, and $^4$He are denoted by
$k/k^\prime$, $q/q^\prime$, and $p/p^\prime$ respectively. 
For large virtual photon's 4-momentum squared, $Q^2=-(k-k')^{2}$, and small squared momentum transfer, $t=(p-p')^{2}$, the DVCS handbag 
diagram can be factorized into two parts \cite{Freund_Collins,Ji_Osborne}. The 
hard part includes the photon-quark interaction and is calculable in 
perturbative QCD. The non-perturbative part is parametrized in terms of GPDs, 
which embed the partonic structure of the hadron. The GPDs depend on the three 
variables $x$, $\xi$, and $t$ as introduced in Fig.~\ref{fig:diags}. The 
parameter $\xi$ relates to the Bjorken variable $x_{B}$: $\xi\approx 
{{x_B}\over{2-x_B}}$, where $x_B=\frac{Q^2}{2M\nu}$, $\nu=k^0-k^{0\prime}$, and 
$M$ is the proton mass. The parameter $x$ is the quark's internal loop momentum 
fraction and cannot be accessed experimentally. We in fact measure 
Compton form factors (CFF) \cite{Guidal:2013rya}, which are
complex amplitudes defined as:

@@

The $^4$He nucleus is an ideal experimental target for DVCS on nuclei, as it is light enough to be detected in such a setting, while it is subject to significant nuclear effects 
\cite{JSeely} and has rather high density. A helium target leads to another 
important advantage, as the number of GPDs defined for a hadron depends on its 
spin. The structure of a spin-zero nucleus, such as $^4$He, is parametrized by 
only one chiral even GPD ($H_{A}(x,\xi,t)$) at leading twist, while 4 GPDs 
arise in the nucleon case. This significantly simplifies the interpretation of 
the data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

The CEBAF large acceptance spectrometer in Hall-B at Jefferson Lab has been
previously used successfully for DVCS measurements on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}.  In order
to extend its capabilities to nuclear DVCS with a fully exclusive final state,
we built a specialized radial time projection chamber (RTPC) for the detection
and identification of low-energy recoiling light nuclei. 
The experiment E08-024 took place in Hall-B at Jefferson Lab in 2009 using the 
nearly 100\% duty factor, longitudinally polarized electron beam (83$\%$ 
polarization) at an energy of 6.064 GeV. The data were collected over 40 days 
using a 6 atm gaseous $^4$He target placed 64 cm upstream of the nominal center 
of CLAS. For DVCS experiments, the CLAS baseline design \cite{Mecking:2003zu} 
is supplemented with an inner calorimeter (IC) and a solenoid magnet. The IC 
extends the photon detection acceptance of CLAS to polar angles as low as 
4$^{\circ}$. The low-energy M\o{}ller electrons produced in the target form a 
very high rate background that is suppressed by a 5-Tesla solenoid magnet 
placed around the target.

@@

The DVCS and well-known Bethe-Heitler (BH) processes, in which the real 
photon is emitted by the incoming or the outgoing lepton, have the same final 
state and are indistinguishable. The amplitude of electroproduction of real
photons includes a sum of the amplitudes of these two processes. The BH 
amplitude is defined by the target form factors, while the DVCS amplitude is 
a combination of the form factors and GPDs. For our kinematics, the cross 
section of real photon electroproduction is dominated by the BH 
contribution, while the DVCS contribution is very small. Its effect is however
enhanced in the observables sensitive to the interference term, {\it e.g.} 
BSA. The three terms entering the cross section calculation,
the square of the BH and DVCS amplitudes and their interference term, depend on the
azimuthal angle $\phi$ between the $(e,e^\prime)$ and $(\gamma^*,^4$He$^\prime)$ planes,
as shown for the nucleon in Ref.~\cite{Belitsky:2001ns} and for the spin-zero targets
in Refs.~\cite{Kirchner:2003wt,Belitsky:2008bz}. Based on this work, the BSA 
of a spin-zero hadron can be expressed as
\begin{equation}
\begin{split}
&A_{LU}(\phi) = \\
&\frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( \Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)},
\end{split}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the $^4$He CFF $\mathcal{H}_{A}$ and depend on the GPD $H_A$.  
Explicit expressions of the purely kinematic factors $\alpha_i$ are written in
Ref.~\cite{Belitsky:2008bz} and are functions of Fourier harmonics in the 
azimuthal angle $\phi$, the nuclear form factor $F_A(t)$, $Q^2$, $x_B$,
$x_{A} = \frac{x_{B}M_{N}}{M_{A}}$, $y=\frac{Q^{2}}{2x_{A}M_{A}E_{beam}}$, and 
$\epsilon = \frac{2x_{A}M_{A}}{\sqrt{Q^{2}}}$, where $M_{A}$($M_{N}$) is the $^4$He 
(nucleon) mass. Using the different $\sin(\phi)$ and $\cos(\phi)$ contributions 
in these factors, it is possible to unequivocally extract both imaginary and real 
parts of $\mathcal{H}_{A}$ by fitting the $A_{LU}(\phi)$ distrubution.

In this work, the azimuthal dependence of the beam-spin asymmetry has been studied 
for a wide range of kinematics ($Q^2$ between 1 and 2.2 GeV$^2$/c$^2$, $x_B$ between
0.1 and 0.3, and $-t$ between 0.07 and 0.16 GeV$^2$/c$^2$). We identified two main
backgrounds to our measurement, accidental coincidences and exclusive coherent
$\pi^0$ production.  The accidentals have particles originating from different events,
and we estimated their contribution to be 4.1\% of our sample. We evaluated this 
contribution by selecting events passing all our cuts but with an electron and 
helium originating from different vertices. The $\pi^0$ production can easily
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low energy in the laboratory frame and remains undetected.  
To estimate the effect of this background, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ measured by our 
experiment. We used this generator together with a GEANT3 simulation of our 
detector to estimate the ratio of the number of $\pi^0$ events where the 
two photons are detected to those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data.  Depending on the kinematics, we found contaminations of 
2 to 4\%.  Studies of systematic uncertainties showed that the main 
contributions come from the choice of DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%), where the values are relative and quoted for $A_{LU}$
at $\phi=90^\circ$.  Added quadratically, the total systematic uncertainty
is about 10\% at $90^\circ$ (or 0.03, absolute), which is significantly smaller
than the statistical uncertainties in all kinematical bins.
